---
title: "R Notebook"
output: html_notebook
---

```{r}
setwd("/Users/ppower1/Desktop/")
```

```{r}
install.packages("readr")
install.packages("dplyr")
install.packages("ggplot2")
install.packages("patchwork")
install.packages("tigris")
install.packages("sf")
install.packages("ivreg") 
```
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(patchwork)
library(tigris)
library(sf)
library(ivreg)
```

**Question #1**:

(a)
```{r}
p <- seq(30000, 80000, length.out = 100)
```

(b)
```{r}
housing_supply <- function(a, p, s, o) {
result <- a * p^s * o
return(result)
}
```

(c)
```{r}
a <- 1.0
o <- 1.0

h_low <- housing_supply(a, p, 0.5, o)
h_high <- housing_supply(a, p, 2.0, o)

# STORE RESULTS IN DATAFRAME
df_plot <- data.frame(
P = rep(p, 2),
H = c(h_low, h_high),
psi = factor(rep(c("ψ = 0.5", "ψ = 2"), each = 100)))

ggplot(df_plot, aes(x = H, y = P, color = psi)) +
geom_line(size = 1.2) +
labs(x = "Housing Supply (Hˢ)", y = "Housing Price",
title = "Housing Supply Function",
color = "") +
theme_minimal()
```

**Question #2**
```{r}
log_h <- seq(0, 25, length.out = 100)

df_log <- data.frame(
  log_H = rep(log_h, 2),
  log_P = c(log_h / 0.5, log_h / 2),
  psi = factor(rep(c("ψ = 0.5", "ψ = 2"), each = 100))
)

ggplot(df_log, aes(x = log_H, y = log_P, color = psi)) +
  geom_line(size = 1.2) +
  labs(x = "log(Hˢ)", y = "log(P)",
       title = "Housing Supply Function (Log Scale)",
       color = "") +
  theme_minimal()
```





**Question #3**

*Any answer is acceptable here -- I just wanted them to think about this question *

**Question #4**
Generous partial credit for reasoning about something. The correct answer is that correlated demand shocks with income growth will have no impact on our estimates.


**Question #5**
The high level idea here is that our estimate will be biased down when supply shocks are more strongly correlated with income growth in low elasticity areas.

**Question 6**

(a): Reading in the Libraries -- done above!

(b)
```{r}
df <- read.csv("msa_analysis_data.csv")
head(df)
```

(c)
```{r}
df <- df %>% filter(!is.na(elasticity), !is.na(g_real_hpi), !is.na(g_housing_units))

```

(d)
```{r}
df$cbsa_code <- as.character(df$cbsa_code)
```

**Question #7**

(a)
```{r}
# Download MSA shapefiles
options(tigris_use_cache = TRUE)
msas <- core_based_statistical_areas(year = 2020)
```


(b)
```{r}
# Merge
merged <- msas %>%
  left_join(df, by = c("GEOID" = "cbsa_code"))
```

(c)
```{r}
ggplot(merged) +
  geom_sf(aes(fill = elasticity), color = NA) +
  scale_fill_viridis_c(name = "Saiz Elasticity", trans = "log", na.value = "white") +
  coord_sf(xlim = c(-125, -66), ylim = c(24, 50)) +
  labs(title = "Housing Supply Elasticity by MSA") +
  theme_minimal()
```

**Question #8**

(a)
```{r}
p1 <- ggplot(df, aes(x = elasticity, y = g_real_hpi)) +
  geom_point() +
  labs(x = "Saiz Elasticity", y = "Real House Price Growth (%)") +
  theme_minimal()

p2 <- ggplot(df, aes(x = elasticity, y = g_housing_units)) +
  geom_point() +
  labs(x = "Saiz Elasticity", y = "Housing Unit Growth (%)") +
  theme_minimal()

p1 + p2
```

(b)
```{r}
ggplot(df, aes(x = g_real_income, y = g_real_hpi, color = factor(elastic_saiz))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Real Income Growth (%)", 
       y = "Real House Price Growth (%)",
       color = "Elastic") +
  scale_color_manual(values = c("0" = "red", "1" = "blue"),
                     labels = c("0" = "Inelastic", "1" = "Elastic")) +
  theme_minimal()
```

**Question #9**

(a)
```{r}
price_reg <- lm(g_real_hpi ~ g_real_income * elastic_saiz, data = df)
summary(price_reg)
```

(b)
```{r}
quantity_reg <- lm(g_housing_units ~ g_real_income * elastic_saiz, data = df)
summary(quantity_reg)
```


**Question 10**: Explore!

Here's what I did. Students are free to do what they want.
```{r}
# --- IV separately by elasticity group ---
df_inelastic <- df %>% filter(elastic_saiz == 0)
df_elastic   <- df %>% filter(elastic_saiz == 1)

iv_inelastic <- ivreg(g_housing_units ~ g_real_hpi | g_real_income, data = df_inelastic)
iv_elastic   <- ivreg(g_housing_units ~ g_real_hpi | g_real_income, data = df_elastic)


# --- Compare elasticities ---
cat("Supply elasticity (inelastic cities):", coef(iv_inelastic)["g_real_hpi"], "\n")
cat("Supply elasticity (elastic cities):  ", coef(iv_elastic)["g_real_hpi"], "\n")
```
```